---
title       : The wine example using K-Means (Clustering technique)
author      : Edurne Alonso Morán
framework   : revealjs        
highlighter : highlight.js  
widgets     : []            
mode        : selfcontained 
knit        : slidify::knit2slides
navigation  : slide
---

## The wine example using K-Means (Clustering technique)

Author : Edurne Alonso Moran

---

## Introduction

A dataset containing 13 chemical measurements (Alcohol, Malic acid, Ash, Alcalinity of ash, Magnesium, Total phenols, Flavanoids, Nonflavanoid phenols, Proanthocyanins, Color intensity, Hue, OD280/OD315 of diluted wines and Proline) on <span style="color:green; font-weight:bold">178 Italian wine samples</span> is analyzed. 

These data are the results of a chemical analysis of wines grown in the same region in Italy but derived from three different cultivars.

--- .class #id

## Methodology

1) A <span style="color:green; font-weight:bold">K-means cluster</span> of the data is performed. This algorithm is the most common partitioning method and aims to partition n observations into <span style="color:green; font-weight:bold">k</span> clusters in which each observation belongs to the cluster with the nearest mean, serving as a prototype of the cluster. This results in a partitioning of the data space into Voronoi cells.

2) Since K-means cluster analysis starts with <span style="color:green; font-weight:bold">k randomly chosen centroids</span>, a different solution can be obtained each time the function is invoked. Indeed, the variables vary in range, so they are <span style="color:green; font-weight:bold">standardized</span> prior to clustering.

3) In the shiny app, the number of clusters (<span style="color:green; font-weight:bold">k</span>) is determined by user and the cluster graph is drawn each time the user changes the number of clusters. 

--- &twocol 

## R code and plot

*** =left
R code
```{r, echo=FALSE}
library(HDclassif);library(cluster);library(fpc)
library(permute);library(lattice);library(vegan)
library(shiny)
```

```{r, echo=TRUE, eval=FALSE}
data(wine)
df<-scale(wine[-1])
c<-kmeans(df,3)
cmd<-cmdscale(dist(df))
groups<-levels(factor(c$cluster))
ordiplot(cmd)
for(i in seq_along(groups)){
  points(cmd[factor(c$cluster)==groups[i],],col=i,pch=16)}
ordispider(cmd,factor(c$cluster),label=TRUE)
ordihull(cmd,factor(c$cluster),lty="dotted")
```

*** =right
R plot (fixed to 3 cluster)
```{r, echo=FALSE, eval=TRUE}
data(wine);df<-scale(wine[-1])
c<-kmeans(df,3)
cmd<-cmdscale(dist(df))
groups<-levels(factor(c$cluster))
options(warn=-1)
ordiplot(cmd)
for(i in seq_along(groups)){points(cmd[factor(c$cluster)==groups[i],],col=i,pch=16)}
ordispider(cmd,factor(c$cluster),label=TRUE)
ordihull(cmd,factor(c$cluster),lty="dotted")
```

---

## My Shiny App

Then, I have implemented the code shown before using an interactive web application - Shiny. This is a new package from RStudio that makes it incredibly easy with R.

The <span style="color:green; font-weight:bold">application</span> designed for this project can be found in the following url:
https://edurnita.shinyapps.io/my_app